Stochastic Approximation Monte Carlo with a Dynamic Update Factor (SAD)

### 2019-03-04 papers/histogram/sad.tex
edits during meeting

\begin{figure}
  \includegraphics[width=\columnwidth]{figs/N50-lndos-comparison}
  \caption{The entropy of a square well fluid with 50 atoms and filling fraction
        0.3 as the green line.  The green hatched area reflects the
        minimum entropy change needed to converge to the true value.
        The light blue area is the quadratic approximation
        for the change in entropy.  The vertical dotted lines represent
        the energy corresponding to $T=1/3$ and $T=\infty$.}
        ~\jpsays{I fixed this to use samc-1e4-s1. Remove comment}
  \label{fig:entropy-cartoon}
\end{figure}

@@

\begin{figure*}
  \includegraphics[width=0.5\textwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-default.pdf}%
\includegraphics[width=0.5\textwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-slow.pdf}
  \caption{(a) The average entropy error for each MC method is shown for $\delta_0 = 0.05\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$ with
               a fixed energy range reference as discussed in the text. The update factor $\gamma_t$ for
               $N=50$ versus iteration for the four different methods: WL, $1/t$-WL, SAMC, and SAD is shown beneath.
           (b) The average entropy error for each MC method is shown for $\delta_0 = 0.005\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$ with
               the same fixed energy range reference used in (a). The update factor $\gamma_t$ for
               $N=50$ versus iteration for the four different methods: WL, $1/t$-WL, SAMC, and SAD is shown beneath.}\label{fig:fast-slow-gamma}
\end{figure*}

~\jpsays{I edited this section}
For the square-well fluid, our simulations treat systems with a particle number
of $50$ and $256$, a well-width of $\lambda = 1.3$ and $\lambda = 1.5$, and a volume
corresponding to a
filling fraction of $\eta = 0.3$ and $\eta = 0.17$ respectively. The simulations
also explore a reasonable displacement distance $\delta_0 = 0.05\sigma$ and
an unreasonably small displacement distance of $0.005\sigma$. The simulations explore the
energy space of the systems at reduced temperatures of
$T_{\text{min}} = 1/3$ and $1.0$.  All simulations lead to the minimum important
energy $E_{\min}$ and maximum entropy energy $E_{\max}$
being calculated (with the exception of the WL methods where both
of these parameters are needed a priori).  In the sections below, we
examine the systems parameters impact on the convergence of the methods.

The SAMC simulations
computed the density of states for the entire range of possible
energies.  The SAD simulations determined the energy range of interest
dynamically as described above, based on a specified $T_{\min}$.  For
the WL and $1/t$-WL simulations, we constrained the simulation to
remain in the energy range corresponding to $T_{\min} < T < \infty$,
as determined by a previous SAMC simulation.  Thus the WL and $1/t$-WL
simulations were given extra information that in practice would not be
available without additional computational effort.

>>>>

\begin{figure}
  \includegraphics[width=\columnwidth]{figs/N50-lndos-comparison}
  \caption{The entropy of a square well fluid with 50 atoms and filling fraction
        0.3 as the green line.  The green hatched area reflects the
        minimum entropy change needed to converge to the true value.
        The light blue area is the quadratic approximation
        for the change in entropy.  The vertical dotted lines represent
        the energy corresponding to $T=1/3$ and $T=\infty$.}
  \label{fig:entropy-cartoon}
\end{figure}

@@

\begin{figure*}
  \includegraphics[width=0.5\textwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-default.pdf}%
\includegraphics[width=0.5\textwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-slow.pdf}
  \caption{(a) The entropy error for each MC method for $N=50$, 
               $\delta_0 = 0.05\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$
               as a function of number of iterations run.  The error is
               averaged over 8 independent simulations, and the best
               and worst simulations for each method are shown as a
               semi-transparent shaded area.
              (b) The update factor $\gamma_t$ versus iteration number
               for the same simulations.
           (c) \davidsays{Make this similar} The average entropy error for each MC method is shown for $\delta_0 = 0.005\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$ with
               the same fixed energy range reference used in (a). The update factor $\gamma_t$ for
               $N=50$ versus iteration for the four different methods: WL, $1/t$-WL, SAMC, and SAD is shown beneath.}\label{fig:fast-slow-gamma}
\end{figure*}

For the square-well fluid, our simulations treat two systems.  The first is
a smaller simulation with a particle number
of 50, a well-width of $\lambda = 1.3$, and a volume
corresponding to a filling fraction of $\eta = 0.3$. The second system
is larger, with a particle number
of 256, a well-width of $\lambda = 1.5$, and a volume
corresponding to a filling fraction of $\eta = 0.17$.
For each system we use a reasonable displacement distance $\delta_0 = 0.05\sigma$,
and for the smaller system we also use an unreasonably small displacement
distance of $0.005\sigma$. The simulations explore the
energy space of the systems with minimum reduced temperatures of
$T_{\text{min}} = 1/3$ for simulations of the smaller system, and $T_{\min}=1$
for the larger system.
All simulations lead to the minimum important
energy $E_{\min}$ and maximum entropy energy $E_{\max}$
being calculated (with the exception of the WL methods where both
of these parameters are needed a priori).  In the sections below, we
examine the systems parameters impact on the convergence of the methods.

The SAMC simulations
computed the density of states for the entire range of possible
energies.  The SAD simulations determined the energy range of interest
dynamically as described above, based on a specified $T_{\min}$.  For
the WL and $1/t$-WL simulations, we constrained the simulation to
remain in the energy range corresponding to $T_{\min} < T < \infty$,
as determined by a previous SAMC simulation.  Thus the WL and $1/t$-WL
simulations were given extra information that in practice would not be
available without additional computational effort, and the SAMC simulations
computed the entropy over the entire range of possible energies, which
required more effort.
